---
title: "Data source / ideas"
author: "Colin Fay"
date: "11/5/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(vegviz)
library(tidyverse)
```

## to read 

+ Position of the Academy of Nutrition and Dietetics: Vegetarian Diets - https://www.eatrightpro.org/-/media/eatrightpro-files/practice/position-and-practice-papers/position-papers/vegetarian-diet.pdf

+ https://adventisthealthstudy.org/studies

+ https://pubs.acs.org/doi/pdf/10.1021/es702969f

+ Role of Life-style and Dietary Habits in Risk of Cancer among Seventh-Day Adventists https://cancerres.aacrjournals.org/content/35/11_Part_2/3513.long


## HOW MANY ADULTS IN THE U.S ARE VEGETARIAN AND VEGAN?

https://www.vrg.org/nutshell/Polls/2016_adults_veg.htm

## Animal vs world pop

```{r}

# population
fls <- tempfile(fileext = ".zip")
  download.file("http://fenixservices.fao.org/faostat/static/bulkdownloads/Population_E_All_Area_Groups.zip", fls)
  unzip(fls,exdir = dirname(fls))
 list.files(dirname(fls)) 
pop <- readr::read_csv(
  file.path(
      dirname(fls), "Population_E_All_Area_Groups_NOFLAG.csv"
    )
  ) %>% 
  filter(Element == "Total Population - Both sexes", Area == "World") %>%
    pivot_longer(
      names_to = "Year", 
      values_to = "Count", 
      cols = tidyselect::matches("Y....$")
    ) %>%
  mutate(Count = Count * 1000, 
         Year = gsub("Y", "", Year), 
         Year = as.numeric(Year))




fls <- tempfile(fileext = ".zip")
  download.file("http://fenixservices.fao.org/faostat/static/bulkdownloads/Production_LivestockPrimary_E_All_Data.zip", fls)
  unzip(fls,exdir = dirname(fls))
df <- readr::read_csv(
    file.path(
      dirname(fls), "Production_LivestockPrimary_E_All_Data.csv"
    )
  )

# CHICKEN

chicken <- df %>%
    select(-matches("Y....F$")) %>%
    select(-matches("Code")) %>%
    mutate(Item = tolower(Item)) %>%
    filter(grepl("meat, chicken", Item)) %>%
    filter( Unit == "1000 Head") %>% 
    select(-Element, - Unit) %>%
    pivot_longer(
      names_to = "Year", 
      values_to = "Head", 
      cols = tidyselect::matches("Y....$")
    ) %>%
    mutate(
      Head = Head * 1000, 
      Year = gsub("Y(....)", "\\1", Year), 
      Year = as.numeric(Year)
    ) %>%
    group_by(Year, Item) %>%
    summarize(
      Head = sum(Head, na.rm = TRUE)
    )


comparison_chicken <- pop %>%
  left_join(
    chicken, by = "Year"
  ) %>%
  select(Year, Count, Head) %>%
  drop_na() %>%
  pivot_longer(
      names_to = "Type", 
      values_to = "Count", 
      cols = c("Count", "Head")
    ) %>%
  mutate(Type = case_when(
    Type == "Count" ~ "Human", 
    TRUE ~ "Chicken"
  )) 

comparison %>%
  ggplot(aes(Year, Count, color = Type)) + 
  geom_line() + 
  ggtheme() +
  scale_y_continuous(labels = scales::comma) 

library(ggplot2)
library(emoGG)

pop %>%
  left_join(
    chicken, by = "Year"
  ) %>%
  select(Year, Count, Head) %>%
  drop_na() %>%
  filter(Year %% 5 == 0) %>%
  mutate(
    one_man_has = Head / Count, 
    #browser(),
    one_man_has_stacked = purrr::map(one_man_has, ~ 1:round(.x))
  ) %>% 
  unnest(cols = c(one_man_has_stacked)) %>%
  ggplot(aes(as.character(Year), one_man_has_stacked)) + 
  geom_emoji(emoji="1f425", size = 0.03) + 
  coord_flip() + 
  ggtheme()

## 

pigs <- df %>%
    select(-matches("Y....F$")) %>%
    select(-matches("Code")) %>%
    mutate(Item = tolower(Item)) %>%
    filter(grepl("meat, total", Item)) %>%
    filter( Unit == "tonnes") %>% 
    select(-Element, - Unit) %>%
    pivot_longer(
      names_to = "Year", 
      values_to = "Tonnes", 
      cols = tidyselect::matches("Y....$")
    ) %>%
    mutate(
      #Head = Head * 1000, 
      Year = gsub("Y(....)", "\\1", Year), 
      Year = as.numeric(Year)
    ) %>%
    group_by(Year, Item) %>%
    summarize(
      Tonnes = sum(Tonnes, na.rm = TRUE)
    ) %>%
  mutate(Tonnes = Tonnes * 1000)


comparison_pigs <- pop %>%
  left_join(
    pigs, by = "Year"
  ) %>%
  select(Year, Count, Tonnes) %>%
  drop_na() %>%
  pivot_longer(
      names_to = "Type", 
      values_to = "Count", 
      cols = c("Count", "Tonnes")
    ) %>%
  mutate(Type = case_when(
    Type == "Count" ~ "Human", 
    TRUE ~ "Meat (kg)"
  )) 

comparison_pigs %>%
  ggplot(aes(Year, Count, color = Type)) + 
  geom_line() + 
  ggtheme() +
  scale_y_continuous(labels = scales::comma) 

library(ggplot2)
library(emoGG)

pop %>%
  left_join(
    pigs, by = "Year"
  ) %>%
  select(Year, Count, Tonnes) %>%
  drop_na() %>%
  filter(Year %% 5 == 0) %>%
  mutate(
    one_man_has = Tonnes / Count, 
    #browser(),
    one_man_has_stacked = purrr::map(one_man_has, ~ 1:round(.x))
  ) %>% 
  unnest(cols = c(one_man_has_stacked)) %>%
  ggplot(aes(as.character(Year), one_man_has_stacked)) + 
  geom_emoji(emoji="1f969", size = 0.05) + 
  coord_flip() + 
  ggtheme()


```


## table Ciqual étendue pour le calcul des apport nutritionnels CALNUT 2020

+  https://ciqual.anses.fr/#/cms/download/node/20

```{r}
library(tidyverse)
fls <- tempfile(fileext = ".xls")
download.file("https://ciqual.anses.fr/cms/sites/default/files/inline-files/Table%20Ciqual%202020_ENG_2020%2007%2007.xls", fls)
df <- readxl::read_excel(fls)
```

```{r}
library(magrittr)
which_one <- "Salt (g/100g)"
df  %>%
  select(alim_nom_eng, {{which_one}}, alim_grp_nom_eng) %>% 
  mutate(`Salt (g/100g)` = as.numeric(`Salt (g/100g)`)) %>%
  #filter(`Cholesterol (mg/100g)` != 100) %>%
  top_n(25,`Salt (g/100g)`) %>%
  # count(alim_grp_nom_eng)
  #select(`Fat (g/100g)`)
  arrange(desc(`Salt (g/100g)`)) %>%
  
  ggplot(aes(reorder(alim_nom_eng, `FA saturated (g/100g)`), `FA saturated (g/100g)`, fill = alim_grp_nom_eng)) +
  geom_col() + 
  coord_flip()

```

## Water footprints of farm animals and animal products (1996-2005)

> Reference: Mekonnen, M.M. & Hoekstra, A.Y. (2011) The green, blue and grey water footprint of crops and derived crop products, Hydrology and Earth System Sciences, 15(5): 1577-1600.

```{r}
library(tidyverse)
fls <- tempfile(fileext = ".zip")
download.file("https://www.waterfootprint.org/media/downloads/Report48-Appendix-V.zip", fls)
untar(fls, exdir = file.path(dirname(fls), "unzipped"))
list.files(file.path(dirname(fls), "unzipped"))
df <- readxl::read_excel(
  list.files(file.path(dirname(fls), "unzipped"), full.names = TRUE),
  2)
source_table <- readxl::read_excel(
  list.files(file.path(dirname(fls), "unzipped"), full.names = TRUE),
  2, skip = 2)
adjusted_table <- source_table %>%
  slice(-1) %>% 
  fill(`Product description (SITC)`) %>%
  select(`Product description (SITC)`, `World Average`) %>% 
  group_by(`Product description (SITC)`) %>%
  summarise(
    `World Average` = sum(as.numeric(`World Average`))
  ) %>%
  mutate(
    `Product description (SITC)` = reorder(`Product description (SITC)`, `World Average`)
  )
```

## Sector by sector: where do global greenhouse gas emissions come from?


> Reference: Mekonnen, M.M. & Hoekstra, A.Y. (2011) The green, blue and grey water footprint of crops and derived crop products, Hydrology and Earth System Sciences, 15(5): 1577-1600.

https://ourworldindata.org/emissions-by-sector

https://github.com/owid/co2-data

```{r}
# df <- read_csv("https://raw.githubusercontent.com/owid/co2-data/master/owid-co2-data.csv")
```

## Methane (CH4) emissions by sector


```{r}
library(tidyverse)
df <- read_csv("data/methane-emissions-by-sector (1).csv")
df %>%
  pivot_longer(names_to = "sector", values_to = "amount", cols = 4:9) %>%
  group_by(Year, sector) %>%
  summarise(
    consumption = sum(amount, na.rm = TRUE)
  ) %>%
  ggplot(aes(Year, consumption, color = sector, group = sector)) + 
  geom_line(size = 2) + 
  scale_colour_viridis_d() + 
  theme_minimal() 

```

## Nitrous oxide (N2O) emissions by sector

```{r}
library(tidyverse)
df <- read_csv("data/nitrous-oxide-emissions-by-sector.csv")
df %>%
  pivot_longer(names_to = "sector", values_to = "amount", cols = 4:9) %>%
  group_by(Year, sector) %>%
  summarise(
    consumption = sum(amount, na.rm = TRUE)
  ) %>%
  ggplot(aes(Year, consumption, color = sector, group = sector)) + 
  geom_line(size = 2) + 
  scale_colour_viridis_d() + 
  theme_minimal()
```

## Per capita N2O: where do our emissions come from?

```{r}
library(tidyverse)
df <- read_csv("data/per-capita-nitrous-oxide-sector.csv")
df %>%
  pivot_longer(names_to = "sector", values_to = "amount", cols = 4:7) %>%
  group_by(Year, sector) %>%
  summarise(
    consumption = sum(amount, na.rm = TRUE)
  ) %>%
  ggplot(aes(Year, consumption, color = sector, group = sector)) + 
  geom_line(size = 2) + 
  scale_colour_viridis_d() + 
  theme_minimal()
```
