---
title: "Saturated Fat by Diet"
author: "Colin Fay"
date: 2020-12-06
slug: "saturated-intake"
categories: ["health"]
tags: ["health", "food", "fat"]
draft: true
params:
  table: intake-diet.csv
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, 
  echo = FALSE, 
  warning = FALSE, 
  message = FALSE
)
```

### `r emo::ji("mag_right")` About the source 

::: {.source}
Scarborough, P., Appleby, P.N., Mizdrak, A. et al. _Dietary greenhouse gas emissions of meat-eaters, fish-eaters, vegetarians and vegans in the UK_. Climatic Change 125, 179–192 (2014). https://doi.org/10.1007/s10584-014-1169-1
::: 

::: {.link}
[link.springer.com/article/10.1007/s10584-014-1169-1](https://link.springer.com/article/10.1007/s10584-014-1169-1)
:::

::: {.quote}
> Assuming an ordered categorisation (high meat → medium meat → low meat → fish-eaters → vegetarians → vegans), there were significant trends towards lower total fat, saturated fat and protein consumption and higher carbohydrate, total sugar, fibre and fruit and vegetables consumption as animal-based food consumption decreased.
:::

::: {.count}
> The diets of 2,041 vegans, 15,751 vegetarians, 8,123 fish-eaters and 29,589 meat-eaters aged 20–79 were assessed using a validated food frequency questionnaire. Comparable GHG emissions parameters were developed for the underlying food codes using a dataset of GHG emissions for 94 food commodities in the UK, with a weighting for the global warming potential of each component gas.
:::

```{r}
library(tidyverse)
library(vegviz)
if (!file.exists(
  params$table
)){
  library(rvest)
  read_html("https://link.springer.com/article/10.1007/s10584-014-1169-1/tables/2") %>%
    html_table() %>%
    purrr::pluck(1) %>% 
    thinkr::clean_names() %>%
    filter(x == "Mean (SD) saturated fat (% energy)") %>%
    pivot_longer(names_to = "Diet", values_to = 'Mean (SD) saturated fat (% energy)', cols = 2:9) %>%
    mutate(
      `Mean (SD) saturated fat (% energy)` = gsub("(.*.).(.) .*", "\\1.\\2", `Mean (SD) saturated fat (% energy)`)
    ) %>%
    filter(Diet != "p_for_trend") %>%
    filter(Diet != "meat_eaters") %>%
    mutate(
      `Mean (SD) saturated fat (% energy)` = as.numeric(`Mean (SD) saturated fat (% energy)`), 
      Diet = gsub("[^_]_a*$", "\\1", Diet)
    ) %>%
    select(-x) %>%
    mutate(
      Diet = case_when(
        Diet == "high_meat_consumer" ~ "High meat-eaters ( ≥ 100 g/day)",
        Diet == "medium_meat_consumer" ~ "Medium meat-eaters (50–99 g/day)",
        Diet == "low_meat_consumer" ~ "Low meat-eaters ( < 50 g/day)",
        Diet == "fish_eaters" ~ "Fish-eaters",
        Diet == "vegetarians" ~ "Vegetarians",
        Diet == "vegans" ~ "Vegans"
      )
    )%>%
    readr::write_csv(params$table)
}


```


### `r emo::ji("chart_with_upwards_trend")` Visualising the data

Visualization of _Table 2 Descriptive statistics_.

```{r}
# DATA MANIP
source_table <- readr::read_csv(params$table)
```


This visualization displays the mean amount of saturated fat eaten every days for a given diet (either `r knitr::combine_words(source_table$Diet, and = "or ")`).
Units are expressed in `% energy`.

```{r saturated-intake, fig.dim = c(12, 8)}
col_vals <- viridis::viridis(6)
names(col_vals) <- c("High meat-eaters ( ≥ 100 g/day)", "Medium meat-eaters (50–99 g/day)", "Low meat-eaters ( < 50 g/day)", "Fish-eaters", "Vegetarians", "Vegans")

source_table  %>%
  ggplot(
    aes(
      reorder(Diet, `Mean (SD) saturated fat (% energy)`),
      `Mean (SD) saturated fat (% energy)`, 
      fill = Diet, 
      label = Diet
    )
  ) + 
  geom_col(alpha = .9) + 
  geom_text(
    aes(y = 0.2,label = Diet, hjust = 0), 
    size = rel(6)
  ) + 
  coord_flip() + 
  scale_fill_viridis_d(option = "D") + 
  guides(
    fill = FALSE
  )  +
  ggtheme() +
  labs(
    title = "The more meat, the more saturated fat every day", 
    subtitle = "From: 'Dietary greenhouse gas emissions of meat-eaters, fish-eaters, vegetarians and vegans in the UK'\nhttps://link.springer.com/article/10.1007/s10584-014-1169-1", 
    x = " ", 
    y = "Mean saturated fat (% energy)	", 
    fill = "Diet type",
    caption = "Via dataviz.vg"
  ) + 
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank())


```

`r details::details(knitr::kable(source_table, format = "html"), summary = "Show the data",lang  = NULL)`

::: {.code}
Code behind this post is available `r build_link(knitr::current_input())`. 
:::
