---
title: "Food and Cholesterol"
author: "Colin Fay"
date: 2020-12-20
slug: "food-and-cholesterol"
categories: ["health"]
tags: ["food", "nutrition"]
draft: true
params:
  table: food-and-cholesterol.csv
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, 
  echo = FALSE, 
  warning = FALSE, 
  message = FALSE
)
```

### `r emo::ji("mag_right")` About the source 

::: {.source}
French Agency for Food, Environmental and Occupational Health & Safety. ANSES-CIQUAL French food composition table version 2020.
::: 

::: {.link}
[ciqual.anses.fr](https://ciqual.anses.fr)
:::

::: {.quote}
>  Individuals should eat as little dietary cholesterol as possible while consuming a healthy eating pattern. In general, foods that are higher in dietary cholesterol, such as fatty meats and high-fat dairy products, are also higher in saturated fats. 
> -- [Dietary Guidelines for Americans](https://health.gov/sites/default/files/2019-09/2015-2020_Dietary_Guidelines.pdf)
:::

::: {.count}
> ANSES-CIQUAL compiles different data sources to prepare the food composition table. This online version of the table provides the list of data sources which were used to produce the values for each pair food – component. ANSES-CIQUAL launches every year a plan to sample and analyse 60 to 80 foods, in collaboration with a subcontractor laboratory. However, this is not sufficient to update all composition data in the table.
:::

```{r}
library(tidyverse)
library(vegviz)
if (!file.exists(
  params$table
)){
  fls <- tempfile(fileext = ".xls")
  download.file("https://ciqual.anses.fr/cms/sites/default/files/inline-files/Table%20Ciqual%202020_ENG_2020%2007%2007.xls", fls)
  readxl::read_excel(fls) %>%
    mutate(`Cholesterol (mg/100g)` = as.numeric(`Cholesterol (mg/100g)`)) %>%
    #filter(`Cholesterol (mg/100g)` != 100) %>%
    #top_n(30, `Cholesterol (mg/100g)`) %>%
    # count(alim_grp_nom_eng)
    #select(`Fat (g/100g)`)
    arrange(desc(`Cholesterol (mg/100g)`)) %>%
    select(alim_nom_eng, `Cholesterol (mg/100g)`, alim_grp_nom_eng) %>%
    # WRITE
    readr::write_csv(params$table)
  unlink(fls)
}

df <- readxl::read_excel(params$table)
df %>%
  mutate(`Salt (g/100g)` = as.numeric(`Salt (g/100g)`)) %>%
  ggplot(aes(x = `Salt (g/100g)`, y = alim_grp_nom_eng, fill = alim_grp_nom_eng)) +
  geom_density_ridges() +
  xlim(0,5) +
  theme_ridges() +
  theme(legend.position = "none")

```


### `r emo::ji("chart_with_upwards_trend")` Visualising the data

```{r}
# DATA MANIP
source_table <- readr::read_csv(params$table)
```


In total, the database contains 3186 aliments. 
We are displaying there the top 30, i.e. around 1% of the dataset. 

```{r fig.dim = c(12,14)}
library(ggridges)
library(ggplot2)

# Diamonds dataset is provided by R natively
#head(diamonds)

# basic example
source_table %>%
  drop_na() %>%
  group_by(alim_grp_nom_eng) %>%
  mutate(m = mean(`Cholesterol (mg/100g)`)) %>%
  ggplot(
    aes(x = `Cholesterol (mg/100g)`, 
        y = reorder(alim_grp_nom_eng, m), 
        fill = alim_grp_nom_eng
    )
  ) +
  geom_density_ridges() +
  xlim(0,150) +
  theme_ridges() +
  guides(fill = FALSE) +
  labs(y = "") +
  theme(legend.position = "none") +
  ggtheme()

```

`r details::details(knitr::kable(source_table, format = "html"), summary = "Show the data",lang  = NULL)`

```{r}
# VIZ
source_table %>%
  drop_na() %>%
  filter(`Cholesterol (mg/100g)` > 150) %>%
  group_by(alim_grp_nom_eng) %>%
  mutate(m = mean(`Cholesterol (mg/100g)`)) %>%
  ggplot(
    aes(x = `Cholesterol (mg/100g)`, 
        y = reorder(alim_grp_nom_eng, m), 
        fill = alim_grp_nom_eng
    )
  ) +
  geom_density_ridges()

  geom_col() +
  coord_flip() +
  #scale_fill_manual(values = viridis::viridis(3)[1:2]) +
  ggtheme() +
  labs(
    title = "The Top 30 Aliments with the Most Cholesterol (mg/100g) \nare Animal-Based",
    subtitle = "From: 'ANSES-CIQUAL French food composition table version 2020' https://ciqual.anses.fr/\nTotal records in the database: 3186",
    x = " ",
    y = "Cholesterol (mg/100g)",
    fill = "Group",
    caption = "Via dataviz.vg"
  )
```

```{r}
# VIZ
source_table %>%
  filter(alim_grp_nom_eng == "fats and oils") %>%
  top_n(10, `Cholesterol (mg/100g)`) %>%
  ggplot(
    aes(reorder(alim_nom_eng, `Cholesterol (mg/100g)`), `Cholesterol (mg/100g)`, fill = alim_grp_nom_eng)
  ) +
  geom_col() +
  coord_flip() +
  scale_fill_manual(values = viridis::viridis(3)[1:2]) +
  ggtheme() +
  labs(
    title = "The Top 30 Aliments with the Most Cholesterol (mg/100g) \nare Animal-Based",
    subtitle = "From: 'ANSES-CIQUAL French food composition table version 2020' https://ciqual.anses.fr/\nTotal records in the database: 3186",
    x = " ",
    y = "Cholesterol (mg/100g)",
    fill = "Group",
    caption = "Via dataviz.vg"
  )
```


::: {.code}
Code behind this post is available `r build_link(knitr::current_input())`. 
:::
