---
title: "Typical Slaughter Age and Natural Life Span of Animals"
author: "Colin Fay"
date: 2020-11-29
slug: "natural-slaughtered-lifespan-animals"
categories: ["industry"]
tags: ["livestock", "production", "industry"]
draft: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, 
  echo = FALSE, 
  warning = FALSE, 
  message = FALSE
)
# READ DATA HERE
library(dplyr)
library(ggplot2)
library(vegviz)

table <- "lifespan.csv"

if (!file.exists(
  table
)){
  # READ DATA HERE
  library(rvest)
  res <- read_html(
    "https://www.farmtransparency.org/kb/abattoirs/48-age-animals-slaughtered"
  )
  
  tbl <- res %>%
    html_table() %>%
    .[[1]]
  
  names(tbl) <- c("animal", "Typical Slaughter Age", "Natural Life Span")
  tbl[-1, ] %>%
    mutate(
      s_n = gsub(
        "(\\d+-)*(\\d+) [a-z]*",
        "\\2",
        `Typical Slaughter Age`,
      ),
      s_factor = case_when(
        grepl("day", `Typical Slaughter Age`) ~ lubridate::days(1),
        grepl("weeks", `Typical Slaughter Age`) ~ lubridate::weeks(1),
        grepl("months", `Typical Slaughter Age`) ~ lubridate::days(30),
        grepl("years", `Typical Slaughter Age`) ~ lubridate::years(1)
      ),
      s_span = as.numeric(s_n) * s_factor,
      `Natural Life Span` = gsub(
        "Up to",
        "",
        `Natural Life Span`,
      ),
      `Natural Life Span` = gsub(
        "\\*",
        "",
        `Natural Life Span`,
      ),
      n_s = gsub(
        "(\\d+-)*(\\d+) [a-z]*",
        "\\2",
        `Natural Life Span`,
      ),
      n_factor = case_when(
        grepl("day", `Natural Life Span`) ~ lubridate::days(1),
        grepl("weeks", `Natural Life Span`) ~ lubridate::weeks(1),
        grepl("months", `Natural Life Span`) ~ lubridate::days(30),
        grepl("years", `Natural Life Span`) ~ lubridate::years(1)
      ),
      n_span = as.numeric(n_s) * n_factor
    ) %>%
    select(
      Animal = animal,
      `Natural Life Span` = n_span,
      `Typical Slaughter Age` = s_span
    ) %>%
    mutate(
      `Natural Life Span` = as.numeric(`Natural Life Span`) / (86400 * 365.25),
      `Typical Slaughter Age` = as.numeric(`Typical Slaughter Age`) / (86400 * 365.25)
    ) %>%
    write.csv(table, row.names = FALSE)
}
source_table <- read.csv(table, stringsAsFactors = FALSE)

```

### `r emo::ji("mag_right")` About the source 

::: {.source}
Farm Transparency Project - Age of animals slaughtered
::: 

::: {.link}
[farmtransparency.org/kb/abattoirs/48-age-animals-slaughtered](https://www.farmtransparency.org/kb/abattoirs/48-age-animals-slaughtered)
:::

::: {.quote}
> Most animals are slaughtered as babies, at a very small portion of their natural life span, regardless of whether they are raised conventionally or in operations that are labelled “humane”, “sustainable”, “natural”, “free range”, “cage free”, “heritage bred”, “grass fed”, “local” or “organic”.
:::


### `r emo::ji("chart_with_upwards_trend")` Visualising the data

Visualization of _Age of animals slaughtered_.

Note: we have rounded numbers using the following patterns: 

+ We took the upper number when availabe (for example, in `6-8 years`, we kept `8`)
+ Months are rounded to 30 days 
+ Every entry was transformed to a number of seconds, then divided by `86400 * 365.25` to get the value in year.


This visualization compare typical slaughter age (on the left, in green) to the natural lifespan of `r nrow(source_table)` animals: `r knitr::combine_words(source_table$Animal, and = "or ")`.

```{r lifespan-natural-slaughter, fig.dim = c(12, 8)}

# VIZ
source_table %>%
  ggplot() +
  geom_segment(
    aes(
      x = Animal,
      xend = Animal,
      y = `Natural.Life.Span`,
      yend = `Typical.Slaughter.Age`
    ),
    color = "grey"
  ) +
  geom_point(
    aes(
      x = Animal,
      y = `Typical.Slaughter.Age`,
      color = "Typical Slaughter Age"
    ),
    size = 5
  ) +
  geom_point(
    aes(x = Animal, y = `Natural.Life.Span`, color = "Natural Life Span"),
    size = 5
  ) +
  scale_color_manual(values = c("#440154FF", "#21908CFF", "#FDE725FF")[1:2]) +
  coord_flip() +
  ggtheme() +
  # scale_fill_manual(values = viridis::viridis(6)[1:2]) +
  labs(
    title = "Typical Slaughter Age and Natural Life Span of 14 Animals",
    subtitle = "Data from farmtransparency.org",
    caption = "Via dataviz.vg",
    x = "",
    y = "Age in years",
    color = ""
  )

```

`r details::details(knitr::kable(source_table, format = "html"), summary = "Show the data",lang  = NULL)`

::: {.code}
Code behind this post is available `r build_link(knitr::current_input())`. 
:::
