---
title: "Greenhouse gas emissions per 2,000 kcal by diet"
author: "Colin Fay"
date: 2020-11-02
slug: "greenhouse-by-diet-2000kcal"
categories: ["environment"]
tags: ["greenhouse", "ecology", "environment"]
params:
  table: greenhousebydiet2000kcal.csv
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, 
  echo = FALSE, 
  warning = FALSE, 
  message = FALSE, 
  cache = TRUE
)
library(tidyverse)
library(vegviz)

if (!file.exists(
  params$table
)){
  library(rvest)
  # Reading and extracting the table
  read_html("https://link.springer.com/article/10.1007/s10584-014-1169-1/tables/3") %>%
    html_table() %>%
    purrr::pluck(1) %>% 
    thinkr::clean_names() %>%
    # Need to take only the rows 3 to 8
    slice(3:8) %>%
    # Turning the col to numeric, then reordering the first column based on adjusted_for_age_and_sex
    mutate(
      adjusted_for_age_and_sex = as.numeric(adjusted_for_age_and_sex), 
      x = reorder(x, adjusted_for_age_and_sex)
    ) %>% 
    # Keeping only the columns we need
    select(Diet = x, `Mean dietary GHG emissions (kgCO2e)` = adjusted_for_age_and_sex) %>%
    readr::write_csv(params$table)
}
```

### `r emo::ji("mag_right")` About the source 

::: {.source}
Scarborough, P., Appleby, P.N., Mizdrak, A. et al. _Dietary greenhouse gas emissions of meat-eaters, fish-eaters, vegetarians and vegans in the UK_. Climatic Change 125, 179–192 (2014). https://doi.org/10.1007/s10584-014-1169-1
::: 

::: {.link}
[link.springer.com/article/10.1007/s10584-014-1169-1](https://link.springer.com/article/10.1007/s10584-014-1169-1)
:::

::: {.quote}
> The production of animal-based foods is associated with higher greenhouse gas (GHG) emissions than plant-based foods.
:::

::: {.count}
> The diets of 2,041 vegans, 15,751 vegetarians, 8,123 fish-eaters and 29,589 meat-eaters aged 20–79 were assessed using a validated food frequency questionnaire. Comparable GHG emissions parameters were developed for the underlying food codes using a dataset of GHG emissions for 94 food commodities in the UK, with a weighting for the global warming potential of each component gas.
:::


### `r emo::ji("chart_with_upwards_trend")` Visualising the data

Visualization of _Table 3 Mean greenhouse gas emissions per 2,000 kcal by diet type and sex_ (we are using the aggregated data, i.e. without the sex variable).

```{r}
adjusted_table <- readr::read_csv(params$table)
```


This visualization displays the mean amount of greenhouse gas emission for 2,000kcal of a given diet (either `r knitr::combine_words(adjusted_table$Diet, and = "or ")`).

```{r greenhousebydiet2000kcal, fig.dim = c(12, 8)}

col_vals <- viridis::viridis(6)
names(col_vals) <- c(
  "High meat-eaters ( ≥ 100 g/day)", 
  "Medium meat-eaters (50–99 g/day)", 
  "Low meat-eaters ( < 50 g/day)",
  "Fish-eaters", 
  "Vegetarians", 
  "Vegans"
)

adjusted_table %>%
  ggplot(
    aes(
      reorder(Diet, `Mean dietary GHG emissions (kgCO2e)`),
      `Mean dietary GHG emissions (kgCO2e)`, 
      fill = Diet, 
      label = Diet
    )
  ) + 
  geom_col(alpha = .9) + 
  geom_text(
    aes(y = 0.2,label = Diet, hjust = 0), 
    size = rel(6)
  ) + 
  coord_flip() + 
  scale_fill_viridis_d(option = "D") + 
  guides(
    fill = FALSE
  )  +
  ggtheme() +
  labs(
    title = "More meat equals more gas emissions per 2,000 kcal of food", 
    subtitle = "From: 'Dietary greenhouse gas emissions of meat-eaters, fish-eaters, vegetarians and vegans in the UK'\nhttps://link.springer.com/article/10.1007/s10584-014-1169-1", 
    x = " ", 
    y = "Mean dietary GHG emissions (kgCO2e) per 2,000 kcal of food", 
    fill = "Diet type",
    caption = "Via dataviz.vg"
  ) + 
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )
```

`r details::details(knitr::kable(adjusted_table, format = "html"), summary = "Show the data",lang  = NULL)`

::: {.code}
Code behind this post is available `r build_link()`. 
:::
