---
title: "Typical Slaughter Age and Natural Life Span of Animals"
author: "Colin Fay"
date: 2020-11-29
slug: "natural-slaughtered-lifespan-animals"
categories: ["industry"]
tags: ["livestock", "production", "industry"]
draft: true
params:
  table: lifespan.csv
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, 
  echo = FALSE, 
  warning = FALSE, 
  message = FALSE
)
library(vegviz)
```

### `r emo::ji("mag_right")` About the source 

::: {.source}
Farm Transparency Project - Age of animals slaughtered
::: 

::: {.link}
[farmtransparency.org/kb/abattoirs/48-age-animals-slaughtered](https://www.farmtransparency.org/kb/abattoirs/48-age-animals-slaughtered)
:::

::: {.quote}
> Most animals are slaughtered as babies, at a very small portion of their natural life span, regardless of whether they are raised conventionally or in operations that are labelled “humane”, “sustainable”, “natural”, “free range”, “cage free”, “heritage bred”, “grass fed”, “local” or “organic”.
:::


```{r}
# READ DATA HERE

library(tidyverse)
library(magrittr)
library(vegviz)
if (!file.exists(
  params$table
)){
  # READ DATA HERE
  library(rvest)
  res <- read_html(
    "https://www.farmtransparency.org/kb/abattoirs/48-age-animals-slaughtered"
  )
  
  tbl <- res %>%
    html_table() %>%
    .[[1]]
  
  names(tbl) <- c("animal", "Typical Slaughter Age", "Natural Life Span")
  tbl[-1, ] %>%
  mutate(
    s_n = gsub(
      "(\\d+-)*(\\d+) [a-z]*",
      "\\2",
      `Typical Slaughter Age`,
    ),
    s_factor = case_when(
      grepl("day", `Typical Slaughter Age`) ~ lubridate::days(1),
      grepl("weeks", `Typical Slaughter Age`) ~ lubridate::weeks(1),
      grepl("months", `Typical Slaughter Age`) ~ lubridate::days(30),
      grepl("years", `Typical Slaughter Age`) ~ lubridate::years(1)
    ),
    s_span = as.numeric(s_n) * s_factor,
    `Natural Life Span` = gsub(
      "Up to",
      "",
      `Natural Life Span`,
    ),
    `Natural Life Span` = gsub(
      "\\*",
      "",
      `Natural Life Span`,
    ),
    n_s = gsub(
      "(\\d+-)*(\\d+) [a-z]*",
      "\\2",
      `Natural Life Span`,
    ),
    n_factor = case_when(
      grepl("day", `Natural Life Span`) ~ lubridate::days(1),
      grepl("weeks", `Natural Life Span`) ~ lubridate::weeks(1),
      grepl("months", `Natural Life Span`) ~ lubridate::days(30),
      grepl("years", `Natural Life Span`) ~ lubridate::years(1)
    ),
    n_span = as.numeric(n_s) * n_factor
  ) %>%
  select(
    Animal = animal,
    `Natural Life Span` = n_span,
    `Typical Slaughter Age` = s_span
  ) %>%
  mutate(
    `Natural Life Span` = as.numeric(`Natural Life Span`) / (86400 * 365.25),
    `Typical Slaughter Age` = as.numeric(`Typical Slaughter Age`) / (86400 * 365.25)
  ) %>%
    readr::write_csv(params$table)
}
```


### `r emo::ji("chart_with_upwards_trend")` Visualising the data

Visualization of _Age of animals slaughtered_.

Note: we have rounded numbers using the following patterns: 

+ We took the upper number when availabe (for example, in `6-8 years`, we kept `8`)
+ Months are rounded to 30 days 
+ Every entry was transformed to a number of seconds, then divided by `86400 * 365.25` to get the value un year.

```{r}
# DATA MANIP
library(tidyverse)
adjusted_table <- readr::read_csv(params$table) 
```


This visualization displays the mean amount of greenhouse gas emission for 2,000kcal of a given diet (either `r knitr::combine_words(adjusted_table$x, and = "or ")`).

```{r fig.dim = c(12, 8)}
# VIZ
adjusted_table %>%
  ggplot() +
  geom_segment(
    aes(
      x = Animal,
      xend = Animal,
      y = `Natural Life Span`,
      yend = `Typical Slaughter Age`
    ),
    color = "grey"
  ) +
  geom_point(
    aes(
      x = Animal,
      y = `Typical Slaughter Age`,
      color = "Typical Slaughter Age"
    ),
    size = 5
  ) +
  geom_point(
    aes(x = Animal, y = `Natural Life Span`, color = "Natural Life Span"),
    size = 5
  ) +
  scale_color_manual(values = viridis::viridis(3)[1:2]) +
  coord_flip() +
  vegviz::ggtheme() +
  # scale_fill_manual(values = viridis::viridis(6)[1:2]) +
  labs(
    title = "Typical Slaughter Age and Natural Life Span of Animals",
    subtitle = "Data from farmtransparency.org",
    caption = "Via dataviz.vg",
    x = "",
    y = "Age in years",
    color = ""
  )

```

`r details::details(knitr::kable(adjusted_table, format = "html"), summary = "Show the data",lang  = NULL)`

::: {.code}
Code behind this post is available `r build_link(knitr::current_input())`. 
:::
