---
title: "Worlwide Meat Production — 1961-2018"
author: "Colin Fay"
date: 2020-11-08
slug: "worldwide-meat-production-1961-2018"
categories: ["industry"]
tags: ["livestock", "production", "industry"]
params:
  table: meat-over-years.csv
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, 
  echo = FALSE, 
  warning = FALSE, 
  message = FALSE
)
library(vegviz)
```

### `r emo::ji("mag_right")` About the source 

::: {.source}
Food and Agriculture Organization of the United Nations - _Livestock Primary_ dataset, 'All Area Groups'.
::: 

_Note: in a first version of this article, we were using the 'All Data' dataset, then switched to the 'All Area Groups' dataset, that already countains the worldwide aggregated data._

::: {.link}
[fao.org/faostat/en/#data/QL](https://www.fao.org/faostat/en/#data/QL)
:::

::: {.count}
> The data on livestock numbers are intended to cover all domestic animals irrespective of their age and the place or purpose of their breeding. Estimates have been made for non-reporting countries as well as for countries reporting incomplete data. However, in certain countries, data for chickens, ducks and turkeys do not yet seem to represent the total number of these birds. Certain other countries give a single figure for all poultry; data for these countries are shown under “Chickens”.
:::

```{r}
library(tidyverse)
library(magrittr)
if (!file.exists(
  params$table
)){
  # READ DATA HERE
  # Reading the data for the production
  fls <- tempfile(fileext = ".zip")
  download.file("http://fenixservices.fao.org/faostat/static/bulkdownloads/Production_LivestockPrimary_E_All_Area_Groups.zip", fls)
  unzip(fls,exdir = dirname(fls))
  
  meat_over_year_all_meats <- readr::read_csv(
    file.path(
      dirname(fls), "Production_LivestockPrimary_E_All_Area_Groups.csv"
    )
  ) %>%
    filter(Area == "World") %>%
    # Removing the columns that matches the pattern Y....F$, and which are encoded
    select(-matches("Y....F$")) %>%
    select(-matches("Code")) %>%
    # Turning the item column to lower for easier manipulation
    mutate(Item = tolower(Item)) %>%
    # keeping only the meat elements, with the production in tonnes
    filter(grepl('meat', Item)) %>%
    filter(Element == "Production", Unit == "tonnes") %>% 
    # Removing these two columns
    select(-Element, - Unit) %>%
    # We turn the wide df to a long one
    pivot_longer(
      names_to = "Year", 
      values_to = "Head", 
      cols = tidyselect::matches("Y....$")
    ) %>%
    # Removing the Y from the year name so that we can use it as a numeric
    mutate(
      Year = gsub("Y(....)", "\\1", Year), 
      Year = as.numeric(Year)
    ) %>%
    # For each Year and item, we do a total of Head (the production in tonnes)
    group_by(Year, Item) %>%
    summarize(
      Head = sum(Head, na.rm = TRUE)
    )
  
  meat_over_year_all_meats %>%
    readr::write_csv(params$table)
}
source_table <- readr::read_csv(params$table)

```


### `r emo::ji("chart_with_upwards_trend")` Visualising the data

Visualization of _Livestock Primary_.
The data is filtered by Item, then the Production in tonnes is kept, grouped by year, and summed. 

```{r}
# DATA MANIP
adjust_and_draw <- function(
  item, 
  color = 1
){
  adjusted_table <- source_table %>%  
    filter(grepl(item, Item))
  
  gg <- adjusted_table %>%
    ggplot(aes(Year, Head)) + 
    geom_col(fill = viridis::viridis(6)[color]) + 
    labs(
      x = " ", 
      y = " ", 
      caption = "Via dataviz.vg",
      subtitle = "From: fao.org/faostat"
    ) +
    scale_y_continuous(labels = scales::comma) + 
    ggtheme()
  list(
    tbl  = adjusted_table, 
    gg = gg
  )
}

```


These visualizations will display the evolution of the total production in tonnes, then of various livestock, from 1961 to 2018.

#### Global meat production in tonnes

```{r meat-over-years, fig.dim = c(12, 6)}
# VIZ
df <- adjust_and_draw("meat, total", color = 2)
df$gg + 
  labs(
    title = "Meat production (in tonnes) — 1961-2018"
  )
```

`r details::details(knitr::kable(df$tbl, format = "html"), summary = "Show the data",lang  = NULL)`

#### Le'ts focus on some species:

```{r}
# VIZ
df <- source_table %>%  
  mutate(Item = tolower(Item)) %>%
  # Keeping our species of interest
  filter(grepl("(duck)|(pig)|(chicken)|(poultry)|(sheep)|(cattle)", Item)) %>%
  # Removing indigenous counts
  filter(!grepl('indigenous', Item)) %>% 
  # Removing the total
  filter(!grepl('total', Item))
```

::: {.animated}

```{r meat-over-years-animated}
library(gganimate)
x <- df %>%
  ggplot(aes(Year, Head, group = Item, color = Item)) + 
  geom_line(size = 1.5) + 
  geom_segment(aes(xend = 2020, yend = Head), linetype = 2, colour = 'grey') + 
  geom_text(aes(x = 2021, label = Item), hjust = 0) +
  xlim(1960, 2030) + 
  guides(color = FALSE) + 
  labs(
    title = "Evolution of the meat production (in tonnes) of 6 species — 1961-2018",
    x = " ", 
    y = " ",
    subtitle = "From: fao.org/faostat", 
    caption = "Via dataviz.vg"
  ) +
  scale_y_continuous(labels = scales::comma) + 
  scale_color_viridis_d(option = "A") +
  ggtheme() + 
  transition_reveal(Year)
animate(x, fps = 5, duration = 15, ref_frame = -1, height = 600, width =1200)

```

:::

`r details::details(knitr::kable(df, format = "html"), summary = "Show the data",lang  = NULL)`

::: {.code}
Code behind this post is available `r build_link(knitr::current_input())`. 
:::
