---
title: "Type of Vegetarian Diet, Body Weight, and Prevalence of Type 2 Diabetes"
author: "Colin Fay"
date: 2021-02-08
slug: "type-diet-type-2-diabet"
categories: ["health"]
tags: ["health", "diabete"]
draft: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, 
  echo = FALSE, 
  warning = FALSE, 
  message = FALSE
)
```

### `r emo::ji("mag_right")` About the source 

::: {.source}
Type of Vegetarian Diet, Body Weight, and Prevalence of Type 2 Diabetes
::: 

::: {.link}
[ncbi.nlm.nih.gov/pmc/articles/PMC2671114/](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC2671114/)
:::

::: {.quote}
> Mean BMI was lowest in vegans and incrementally higher in lacto-ovo vegetarians, pesco-vegetarians, semi-vegetarians, and nonvegetarians. Prevalence of type 2 diabetes increased from 2.9% in vegans to 7.6% in nonvegetarians; the prevalence was intermediate in participants consuming lacto-ovo (3.2%), pesco (4.8%), or semi-vegetarian (6.1%) diets. After adjustment (...) vegans, lacto-ovo vegetarians, pesco-vegetarians, and semi-vegetarians had a lower risk of type 2 diabetes than nonvegetarians.
:::

::: {.count}
> The study population comprised 22,434 men and 38,469 women who participated in the Adventist Health Study-2 conducted in 2002–2006. We collected self-reported demographic, anthropometric, medical history, and lifestyle data from Seventh-Day Adventist church members across North America. The type of vegetarian diet was categorized based on a food-frequency questionnaire. We calculated odds ratios (ORs) and 95% CIs using multivariate-adjusted logistic regression.
:::

```{r}
library(tidyverse)
library(vegviz)
table <- paste0("type-diet-type-2-diabet", ".csv")
if (!file.exists(
  table
)){
  library(rvest)
  # READ DATA HERE
   tbls <- read_html("https://www.ncbi.nlm.nih.gov/pmc/articles/PMC2671114/") %>%
    # MANIPULATE
    html_table() 
   tbls[[2]] %>%
     .[2, 2:6] %>%
     pivot_longer(
       cols = everything(), 
       names_to = "Diet", 
       values_to = "Type 2 diabetes"
     ) %>%
    # WRITE
    readr::write_csv(table)
}
source_table <- readr::read_csv(table)
```


### `r emo::ji("chart_with_upwards_trend")` Visualising the data

Comparison of the diabetes prevalence among 5 diets (either `r knitr::combine_words(source_table$name, and = "or ")`).

```{r fig.dim = c(12, 8)}
# VIZ
source_table %>%
    ggplot(
    aes(
      reorder(Diet, `Type 2 diabetes`),
      `Type 2 diabetes`, 
      fill = Diet,  
      label = Diet
    )
  ) + 
  geom_col(alpha = .9) + 
  geom_text(
    aes(y = 0.2,label = `Diet`, hjust = 0), 
    size = rel(6)
  ) + 
  coord_flip() + 
  scale_fill_viridis_d(option = "D") + 
  guides(
    fill = FALSE
  )  +
  ggtheme() +
  labs(
    title = "Vegan and vegetarian diets are associated with a reduction in risk of type 2 diabetes", 
    subtitle = "From: 'Type of Vegetarian Diet, Body Weight, and Prevalence of Type 2 Diabetes'\nhttps://www.ncbi.nlm.nih.gov/pmc/articles/PMC2671114/", 
    x = " ", 
    y = "Prevalence of type 2 diabetes", 
    fill = "Dietary pattern",
    caption = "Via dataviz.vg"
  ) + 
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank())
```

`r details::details(knitr::kable(source_table, format = "html"), summary = "Show the data",lang  = NULL)`

::: {.code}
Code behind this post is available `r build_link(knitr::current_input())`. 
:::
